---
title: "HW11"
author: "Tural Ismayilov & Polad Mahmudov & Mansur Alizada"
date: "December 8, 2017"
output: pdf_document
---

**Team named "GARABAGH" members:**  
Tural Ismayilov (Software Enigeenring)
Mansur Alizada (Computer Science)
Polad Mahmudov (Software Engineering)

**We are working on the project named House Prices: Advanced Regression Techniques on the Kaggle**  

**link: **  
https://www.kaggle.com/c/house-prices-advanced-regression-techniques


**Bitbucket link:**  
https://turalismayilov@bitbucket.org/garabagh/dm.git

#EX1  
#Business understanding  
##Identifying your business goals  
###Background  
Our customers, want to sell a house and they do not know the price which they can take - it can't be too low or too high. To find house
price they usually try to find similar properties in their neighborhood and based on gathered data they try to assess their house price.
But this is not always good. (Guessing the price by looking at similar house may be useful, but there can be certain details that have big
influence on the price, but are not considered therefore may make difficult the selling proccess.) Although that method can work, but
there is better way for predicting house price.

###Bussiness goals:
Increase the gain of seller.

###Bussiness success criteria:
Sales increasing, both the seller and buyer stay satisfied with the price of house.

##Assessing your situation
###Inventory of resources
With 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa, with its description data is available.

###Requirements, assumptions, and constraints
Achieving top 20% result with RMSE in Kaggle is required. 

###Risks and contingencies
If an Internet outage in our dormitory could pose a problem, perhaps our contingency could be to work at university until the outage has
ended.
If the prediction is not well enough, our contingency is to approach our instructor for help.

###Terminology
**Bias** - Positive values of bias indicate the model tends to overestimate the price (on average) while negative values indicate the
model tends to underestimate price.  
**Maimum Deviation** - It identifies the worst prediction tey made in the validation data set.  
**Training set** - this is a set of examples used to fit the parameters of the mode.  
**Validation set** - The fitted model is used to predict the responses fot the observations in a second dataset called validation dataset.  
**Test set** - is a dataset used to provude an unbiased evaluation of final model fit on the training dataset.  
**Root Mean Square Error** - used to obtain the coefficient estimates from the original dataset.  
**Mean Absolute Deviation** - Average error regardless of sign.  

###Costs and benefits
The cost is the loss in gain when the house is sold with underestimated price, and gain happens when it's sold with enough gain.

##Defining your data-mining goals
###Data-mining goals
Data mining goal is to predict house price for its indicators.

###Data-mining success criteria
Prediction with RMSE less than 0.11979.

EX2
#Data understanding
##Gathering data
###Outline data requirements
We need previous sales of house and its details to fit regression model on them for prediction.

###Verify data availability
Approximately 6 percent of elements are NA, so in that case, we will use mice for handling missing data.
```{r}
train <- read.csv('train.csv')
(sum(is.na(train)) / (nrow(train)*ncol(train))) * 100
```

###Define selection criteria
The data will be taken from the link provided above.
```{r}
train <- read.csv('train.csv') 
str(train)
```

###Describing data
```{r}
cat(readLines('data description.txt'), sep = '\n')
```

###Exploring data
```{r}
summary(train)
```

###Verifying data quality
There are some missing data, we have decided to solve this problem by using mice. This is the initial plan, if the mice method does not
work properly we will narrow data size. 

#EX3
##Setting up and planning your project
###Create a project repository either in GitHub or Bitbucket.  

https://www.kaggle.com/c/house-prices-advanced-regression-techniques  

###Register you project by adding a new entry into the List of projects  
New slide was added also to the project repository and the link is:  
https://docs.google.com/presentation/d/1veA_WQcfRRx7hQnE8qklmsLYceWzQGrPHrieSEWqcaI/edit#slide=id.g2a4c3a4a7d_6_0  

###Make a detailed plan of your project with a list of tasks. Specify how many hours each team member is going to contribute to each task

**Acquiring the data and create our environment  
  - Downloading the data  
  - Doing some plotting  
Exploring the data and engineer Features  
  - Working with Numeric Features  
	- Displaying the correlation between columns  
	- Creating a pivot table to further investigate the relationships  
	- Handling null values  
	- Wrangling the non-numeric Features  
	- Transforming features  
Building a linear model  
Begining Linear regression modelling  
Evaluating the performance and visualize results  
Trying to improve the model  
Making a submission  
	- Creating csv file with ids and predictions according to the sample format  
Submitting our results**

For the initial calculations each team member will spend approximately 18 hours.  

###Add the results from business understanding, data understanding and planning to your project repository. Report the links to where these results are listed.

It was added to project repository.

###Prepare to pitch your project at the practice session